---
title       : "Investigating Climate Hindcasts and Predictions of General Circulation Models"
subtitle    : "Temperature Predictions at https://nicolegoebel.shinyapps.io/temperature_map_prediction/"
author      : Nicole Goebel
job         : Coursera Data Science Specialization - Developing Data Products
framework   : html5slides   # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js   # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Goal: to Investigate Hindcasts and Predictions in Temperature of General Circulation Models (GCMs)

- The World Bank hosts predictive model outputs of changes in temperature and precipitation under different climate change scenarios and time periods, avalable in the [World Bank Climate Change Knowledge Portal](http://sdwebx.worldbank.org/climateportal/index.cfm). 
- Ensembles of different GCMs hindcast (`past` scenario) and predict (`a2` or `b1` scenarios) changes in air temperature and precipitation world-wide.
- These GCMs are the same models used in the [Intergovernmental Panel on Climate Change 4th Asessment Report](http://www.ipcc.ch/publications_and_data/publications_ipcc_fourth_assessment_report_synthesis_report.htm).
- The application presented here allows the user to investigate these temperature hindcasts and predictions across the globe for two different scenarios and/or years.

--- .class #id 

## Obtaining Model Data

- Model data and functions to manipulate data are available in the `rWBclimate` R package.
- This package can be found on [GitHub](http://github.com/ropensci/rWBclimate) and is described in the R [vignette](http://cran.r-project.org/web/packages/rWBclimate/vignettes/rWBclimate.pdf) and [pdf](http://cran.r-project.org/web/packages/rWBclimate/rWBclimate.pdf).
- The extraction of annually averaged temperature data for all countries is demonstrated here:

```{r}
library(rWBclimate)
world <- c(NoAm_country, SoAm_country, Eur_country, Asia_country, 
           Africa_country, Oceana_country)
world<-world[! world %in% c("UMI")]  #remove "UMI"
world_dat <- get_ensemble_temp(world, "annualavg", 
           start=1920, end=1939) #select one timeframe to plot
world_dat$data <- as.numeric(as.character(world_dat$data)) #numeric
world_dat<-subset(world_dat,world_dat$percentile==50) #use median 
world_dat$years=paste(world_dat$fromYear,world_dat$toYear, sep="-")
world_dat<-subset(world_dat, select=-c(percentile, fromYear, toYear)) 
```

--- .class #id 

## Plotting Model Data
- The climate_map() function is available in the `rWBclimate` package and is used to plot the data on a world map (below). Path is set for caching kml files, which contain location data.
```{r, echo=FALSE}
library(ggplot2)
```
```{r, fig.show='asis', results='hide', fig.height=5, fig.width=8, fig.align='center'}
options(kmlpath = "/Users/nicolegoebel/Dropbox/Courses/Coursera_Data_Science_2014/DevelopingDataProducts/tempPredict_slidify")
world_map_df <- create_map_df(world)
climate_map(world_map_df,world_dat,return_map = T) + scale_fill_gradient2(name="Temperature", limits=c(-20, 34), low="blue", mid="white", high = "red")
```

--- .class #id 

## Instructions for App: Available Selections

- The user makes two sets of selections. Each selection plots a global temperature map.
- For each map the user chooses:
    1. one scenario: 
      - `past` (years prior to 2000) 
      - `a1` (a future scenario where greenhouse gases reach 850 ppm by 2100)
      - `b1` (a future scenario where greenhouse gases reach 550 ppm by 2100)
    2. a two-decade period over which an annual average is predicted, which is conditional on the selected scenario:
      - past scenario: "1920-1939", "1940-1959", "1960-1979", "1980-1999"
      - future scenarios: "2020-2039", "2040-2059","2060-2079", "2080-2099"
- The result is 3 maps (e.g., map in previous slide) showing the two selected hindcasts and/or predictions and a third map showing the difference between the two.
- check it out at: https://nicolegoebel.shinyapps.io/temperature_map_prediction/
